---
title: "HappyMarkdown"
author: "Group 16"
date: "20/06/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA)
```

```{r libraries}
library(tidyverse)
library(moderndive)
library(skimr)
library(kableExtra)
library(gridExtra)
library(dplyr)
library(ggplot2)
library(infer)
library(GGally)
library(plotly)
```

```{r data, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
#setwd(dir = "D:\\STUDY\\GLASGOW\\Semester-2\\Data Analysis Skills\\Gruopwork\\Groupwork - 1 poster")
happy <- read_csv("happiness.csv")
happy <- as_tibble(happy)
```

This code adds a column to the data, so each country is either Africa or Rest of the World.

```{r addcol}

happy <- happy %>%
  mutate(African = ifelse(happy$`Regional indicator` == "Sub-Saharan Africa" | happy$`Regional indicator` == "Middle East and North Africa",
                "Africa","World"))

```

# Introduction {#sec:Intro}

The World Happiness Dataset is the iconic statistical data of global happiness. It contains six critical factors: economic production, social support, life expectancy, freedom, absence of corruption, and generosity. To find which factors are most influential to the ladder score, we fitted a linear regression model with all relevant explanatory variables. Noting that Africa’s ladder score seems the smallest amongst all countries, we used bootstrapping to find the 95th percentile of the difference in ladder score between Africa and the rest of the world and fit the regression model for these two cases.

# Explanatory analysis {#sec:Explan}

```{r summ}
happy %>%
  group_by(`Regional indicator`) %>%
  summarise(n=n(),  Mean=round(mean(`Ladder score`),2), SD=round(sd(`Ladder score`),2), FQ=quantile(`Ladder score`, 0.25), TQ=quantile(`Ladder score`, 0.75),
            min = min(`Ladder score`), max=max(`Ladder score`),
            median=median(`Ladder score`)) %>%
  kable(col.names = c("Region","Number", "Mean", "SD", "1st Q", "3rd Q", "Min.", "Max","Median"),        
        caption = '\\label{tab:summary} Summary statistics') %>%
  kable_styling(font_size = 10, latex_options = "hold_position")

```


Also summary data for Africa and Rest of the World

```{r sumafrica}
#summary stats for AFRICA and REST OF WORLD
happy %>%
  group_by(African) %>%
  summarise(n=n(),  Mean=round(mean(`Ladder score`),2), SD=round(sd(`Ladder score`),2), FQ=quantile(`Ladder score`, 0.25),
            TQ=quantile(`Ladder score`, 0.75),
            min = min(`Ladder score`), max=max(`Ladder score`),
            median=median(`Ladder score`))  %>%
  kable(col.names = c("Region","Number", "Mean", "SD", "1st Q", "3rd Q", "Min.", "Max","Median"),        
        caption = '\\label{tab:summary} Summary statistics') %>%
  kable_styling(font_size = 10, latex_options = "hold_position")


```


This code uses bootstrapping to find a 95% confidence interval for the mean ladder score of all the data


```{r bootstrap}
bootstrap_distribution_mean <- happy %>% 
  specify(response = `Ladder score`) %>% 
  generate(reps = 1000) %>% 
  calculate(stat = "mean")

bootstrap_distribution_mean %>% 
  get_ci(level = 0.95, type = "percentile")

percentile_ci <- bootstrap_distribution_mean %>% 
  get_ci()

bootstrap_distribution_mean %>% 
  visualize(endpoints = percentile_ci, direction = "between")


```

This next section of code produces a boxplot of ladder score, using region as a factor for all regions.

```{r box, eval = TRUE, out.width = '80%', fig.align = "center", fig.cap = "\\label{fig:scat} Ladder score by region.", fig.pos = "H"}
ggplot(data = happy, mapping = aes(x = factor(`Regional indicator`), 
                                         y = `Ladder score`)) + geom_boxplot(fill = "slateblue") + 
  labs(x = "Region", y = "Score", title = "Happiness score")  + 
  scale_x_discrete(labels = unique(happy$`Regional indicator`))

```


Now same boxplot but just for Africa and rest of the World

```{r boxaf, eval = TRUE, out.width = '80%', fig.align = "center", fig.cap = "\\label{fig:bplot} Ladder score by Africa", fig.pos = "H"}
ggplot(data = happy, mapping = aes(x = factor(African), 
                                   y = `Ladder score`)) + geom_boxplot(fill = "slateblue") + 
  labs(x = "Region", y = "Score", title = "Happiness score")  + 
  scale_x_discrete(labels = c("Africa","Rest of World"))

```


Using bootstrapping (week 7), we can find a 95th percentile of the difference in ladder score between Africa and rest of World.  
This shows that the score is between 1.65 and 1.11 lower in Africa than rest of World.

```{r boot}

bootstrap_distribution <- happy %>%
  specify(`Ladder score`~ African) %>%
  generate(reps = 1000) %>%
  calculate(stat = "diff in means", order = c("Africa", "World"))
percentile_ci <- bootstrap_distribution %>%
  get_ci()

percentile_ci

```

# Linear models {#sec:models}

First we fit the full model includes six variables.

$$\widehat{\mbox{Ladder score}} = \widehat\alpha + \widehat\beta_{\mbox{1}} \cdot \mbox{GDP}
+\widehat\beta_{\mbox{2}} \cdot \mbox{Social support} + \widehat\beta_{\mbox{3}} \cdot \mbox{life expectancy} + \widehat\beta_{\mbox{4}} \cdot \mbox{Freedom} + \widehat\beta_{\mbox{5}} \cdot \mbox{Generosity} + \widehat\beta_{\mbox{6}} \cdot \mbox{corruption}, $$

where 

$\widehat\alpha$ is the intercept;

$\widehat\beta_{\mbox{i}}$ is the coefficeint for each variable (i=1,...,6);

```{r full_model}
full_model<-lm(`Ladder score`~`Logged GDP per capita`+`Social support`+`Healthy life expectancy`+`Freedom to make life choices`+`Generosity`+`Perceptions of corruption`,data = happy)
get_regression_table(full_model)%>%
  dplyr::select(term,estimate) %>%  
  kable(caption = '\\label{tab:reg} Estimates of the parameters from the fitted linear regression model.') %>%
  kable_styling(latex_options = 'HOLD_position')

summary(full_model)
```

It shows that variables "Social support" and "Freedom to make life choice" are most significant, then we check the correlationship between these six variables. 

```{r correlation}
ggpairs(happy[,c(3,7:12)])
```

Although "Healthy life expectancy" has the strongest correlation with "Ladder score", this variable also has the high correlation with other two variables (Logged GDP per capita and Social support). 

Finally, we decide to use two variables, Social support and Freedom to make life choices to fit our model.

$$\widehat{\mbox{Ladder score}} = \widehat\alpha 
+\widehat\beta_{\mbox{1}} \cdot \mbox{Social support} + \widehat\beta_{\mbox{2}} \cdot \mbox{Freedom}, $$

where 

$\widehat\alpha$ is the intercept;

$\widehat\beta_{\mbox{i}}$ is the coefficeint for each variable (i=1,2);

```{r final_model}
final_model<-lm(`Ladder score`~`Social support`+`Freedom to make life choices`,data = happy)
get_regression_table(final_model)%>%
  dplyr::select(term,estimate) %>%  
  kable(caption = '\\label{tab:reg} Estimates of the parameters from the final fitted linear regression model.') %>%
  kable_styling(latex_options = 'HOLD_position')
```

Our final model shows as follows,

$$\widehat{\mbox{Ladder score}} = -1.439
+5.649 \cdot \mbox{Social support} + 2.993 \cdot \mbox{Freedom}, $$
